---
title: "Tutorial on the eenvlp R package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

This tutorial introduces the `eenvlp` R package, designed for multivariate regression: 
$$
  y = B x + \epsilon,
$$
where:

* $y\in\mathbb R^r$ is the vector of response variables,

* $x\in\mathbb R^p$ is the vector of predictor variables,

* $B\in\mathbb R^{r\times p}$ is the unknown coefficient matrix, and

* $\epsilon$ represents the error term.  

The `eenvlp` package is useful in both low- and high-dimensional scenarios, whether the number of predictors $p$ is smaller or larger than the number of observations $n$.

In this tutorial, we will use the `eenvlp` package to analyze the cereal dataset and predict its chemical properties based on spectroscopy data.

To get started, load the `eenvlp` package into your R session:
```{r, message=FALSE, warning=FALSE}
library(eenvlp)
```

## Data Overview

We will use the `cereal` dataset, which includes measurements of two chemical properties (starch and ash) along with infrared spectroscopy predictors.

#### Load and Inspect the Dataset
First, let's load the dataset and inspect its structure to understand the dimensions of our data:
```{r, hold=TRUE}
data(cereal) # load data
n <- nrow(cereal$x) # Number of observations
r <- ncol(cereal$y) # Number of response variables
p <- ncol(cereal$x) # Number of predictors
```
```{r, echo=FALSE}
cat("n", n, "\n")
cat("r:", r, "\n")
cat("p:", p, "\n")
```

The `cereal` dataset consists of 15 cereal samples, each with 145 infrared spectra (predictors) and 2 chemical properties (responses). This high-dimensional setting is ideal for demonstrating the capabilities of the `eenvlp` package.

#### Goal -- Prediction Task
Our goal is to use the enhanced envelope estimator to predict the chemical properties for the first cereal sample. We will split the dataset into a training set and a test set:
```{r}
# Define training and test sets
X_test <- cereal$x[1,] # The first observation as the test set
Y_test <- cereal$y[1,]
X_train <- cereal$x[-1,] # The remaining observations as the training set
Y_train <- cereal$y[-1,]
```

## Model Fitting -- Enhanced Envelope Estimator
To fit the enhanced envelope estimator, we will use cross-validation to select the optimal tuning parameters, $\lambda$ (regularization parameter) and $u$ (dimension of the envelope subspace).

#### Define Sequences for Tuning Parameters
We begin by defining sequences for $\lambda$ and $u$:
```{r}
lambda.seq = 10^(seq(3, -3, length.out=20)) # Sequence of lambda values
u.seq = c(1:r) # Sequence of u values
cv.lamb.env = matrix(NA,length(lambda.seq),length(u.seq))
```

#### Cross-Validation for Parameter Selection
Next, we perform cross-validation to find the optimal values for $\lambda$ and $u$: 

```{r, hold=TRUE}
set.seed(11111) # Set a seed for reproducibility
  for(i in 1:length(lambda.seq)){
    for(j in 1:length(u.seq)){
      cv.lamb.env[i,j] <- cv_eenvlp(X_train, Y_train, u=u.seq[j], lamb=lambda.seq[i], m=10, nperm=1)
    }
  }  

# Identify the optimal parameters
chosen <- which(cv.lamb.env == min(cv.lamb.env), arr.ind=TRUE)
chosen.lamb <- lambda.seq[chosen[1]]
chosen.u <- u.seq[chosen[2]]
```
```{r, echo=FALSE}
cat("Optimal lambda:", chosen.lamb, "\n")
cat("Optimal u:", chosen.u, "\n")
```

#### Fit the `eenvlp` Model Using Optimal Parameters

With the optimal parameters determined, we fit the `eenvlp` model:

```{r}
fit_eenvlp <- eenvlp(X_train, Y_train, u = chosen.u, lamb = chosen.lamb)
```

## Prediction on Test Data
We use the fitted model to predict the chemical properties for the test observation:
```{r}
predict_eenvlp <- pred_eenvlp(fit_eenvlp, X_test)
```

#### Calculate Prediction Error 
Finally, we calculate the prediction error to evaluate the model's performance:
```{r, hold=TRUE}
prediction_error <- (predict_eenvlp - Y_test)^2
```
```{r, echo=FALSE}
cat("Prediction error:", prediction_error, "\n")
```
